Clustering time series based on probability distributions across temporal granularities

### 2021-11-15 main.Rmd
Figure 9 update

The $24$ prototypes <!--and the data subset consisting of $353$ customers--> are clustered using the methodology described in Section \ref{sec:methodology} and results are reported below. In the following plots, the median is shown by a line, and the shaded region shows the area between the  $25^{th}$ and $75^{th}$ percentiles.<!--All customers with the same color represent the same clustered groups. --> Groups by JS-based distances and wpd-based distances are colored differently as they represent different groupings. The plotting scales are not displayed since we want to emphasize comparable patterns rather than scales. The idea is that a customer in a cluster may have low total energy usage, but their behavior may be quite similar to a customer with high usage with respect to distributional pattern or significance across cyclic granularities.

@@

For clustering based on JS-based (JS-NQT) distances, we chose the optimal number of clusters using (@Hennig2014-ah) as $5$. The groupings are shown in Figure \ref{fig:hod-ind-group-png}. <!--Customers with the same color represent the same clustered groups.--> Our methodology is useful for grouping similar distributions over `hod` and `moy` and they are placed closely for easy comparison. Few groups have mixed patterns across `hod` and `moy`, but few have all customers in the group having a similar profile. Figure \ref{fig:hod-ind-group-png} shows the summarized distributions across $5$ groups and assists us in characterizing each cluster. It shows Groups $2$ and $4$ have `hod` pattern with a typical morning and evening peak, whereas groups $1$, $3$, and $5$ show a `moy` pattern with higher usage in winter months. Differences in Weekend/Weekday between groups are not discernible, implying that it may not be a relevant variable in distinguishing various clusters unless maybe conditioned by `moy` or `hod`. It may be interesting to compare these two plots to verify if the summarized distributions across groups correctly characterised the groupings. If it has, then the majority of the group's members should share a similar profile.

@@

```{r combined4and5, fig.cap = "For $k = 4$ (a) and $k=5$ (b), the distribution of electricity demand across groups over hod, moy, and wnwd is shown. groupA-2, groupA-3, and groupA-4 profiles correspond to groupB-2, groupB-3, and groupB-5, respectively. groupA-1 is split into groupB-1 and groupB-4, each of which has a distinct shape across moy and wnwd. groupB-4 (id: 16-20) is characterized by higher energy consumption in the first few months of the year and more variation in weekend usage. groupB-1 (id:1-3) is distinguished by higher consumption in the middle of the year (winter months) and similar weekday-weekend inter-quartile range. When $k = 4$ is used, these two groups merge to form groupA-1, which has a moy profile of higher usage in both the beginning and middle of the year, which is not representative of the individuals in the group."}
knitr::include_graphics("figs/combined4and5.png")
```

>>>>

The $24$ prototypes <!--and the data subset consisting of $353$ customers--> are clustered using the methodology described in Section \ref{sec:methodology} and results are reported below. In the following plots, the median is shown by a line, and the shaded region shows the area between the  $25^{th}$ and $75^{th}$ percentiles.<!--All customers with the same color represent the same clustered groups. --> Groups by JS-based distances and wpd-based distances are colored differently as they represent different groupings.
While plotting the clusters, RS-based scaling is applied to each customer, so that no customer dominate the shape of the summarized clusters because of magnitude. The plotting scales are not displayed since we want to emphasize comparable patterns rather than scales. The idea is that a customer in a cluster may have low total energy usage, but their behavior may be quite similar to a customer with high usage with respect to distributional pattern or significance across cyclic granularities.

@@

The $sindex$ is plotted against the number of clusters for clustering based on JS-based (JS-NQT) distances in \ref{fig:opt-cluster-tsne-jsd}. The $sindex$ falls from around $0.56$ when $k=3$ to $0.54$ when $k$ increases. In our study, we use $k=4, 5$. Figure \ref{fig:combined4and5} depicts the summarised distributions across $4$ and  $5$ clusters in (a) and (b) respectively, and assists in characterizing each cluster. A-2, A-3, and A-4 have profiles that correspond to groups B-2, B-3, and B-5, respectively. B-2 (id: 4-9 in Figure \ref{fig:hod-ind-group-png}) and B-5 (id: 21-24) have a 'hod' pattern with a typical morning and evening peak, whereas the other groups have no distinct pattern or high variability throughout the day. A-1 is subdivided into B-1 and B-4, each of which has a distinct shape across 'moy' and 'wnwd'. B-4 (id: 16-20) is distinguished by higher energy consumption in the first few months of the year and greater variability in weekend usage. B-1 (id: 1-3) has higher consumption in the middle of the year (winter months) and a similar weekday-weekend inter-quartile range. When $k = 4$ is used, these two groups merge to form A-1, which has a moy profile of higher usage in both the beginning and middle of the year, which is not representative of the individuals. It may be worthwhile to compare Figures \ref{fig:opt-cluster-tsne-jsd} and \ref{fig:hod-ind-group-png} to see if the summarized distributions across groups accurately characterized the groupings. If it has, then the majority of the members of the group should have a similar profile.

<!-- Our methodology is useful for grouping similar distributions over `hod` and `moy` and they are placed closely for easy comparison. Few groups have mixed patterns across `hod` and `moy`, but few have all customers in the group having a similar profile.  -->

@@

```{r combined4and5, fig.cap = "For $k = 4$ (a) and $k=5$ (b), the distribution of electricity demand across groups over hod, moy, and wnwd is shown. Groups A-2, A-3, and A-4 profiles correspond to Groups B-2, B-3, and B-5, respectively. A-1 is split into B-1 and B-4, each of which has a distinct shape across moy and wnwd. B-4 (id: 16-20) is characterized by higher energy consumption in the first few months of the year and more variation in weekend usage. B-1 (id:1-3) is distinguished by higher consumption in the middle of the year (winter months) and similar weekday-weekend inter-quartile range. When $k = 4$ is used, these two groups merge to form A-1, which has a moy profile of higher usage in both the beginning and middle of the year, which is not representative of the individuals in the group."}
knitr::include_graphics("figs/combined4and5.png")
```
